---
title: "Analyzing Sleep Through Smart Technology: A Case Study of Will Foote’s Fitbit Data"
subtitle: "[CODE FILE]"
author: "William Foote & Kyle Fang"
date: "6/12/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading sleep data

```{r}
library(readxl)
library(chron)
sleep <- read_excel("combined_data_final.xlsx")
class(sleep$Date)
```

# Clean start and end times for sleep

## Need get_miltime function

```{r}
get_miltime <- function (x) {
  if (grepl("-", x)) {
    x <- gsub("-", "", x)
    x <- unlist(strsplit(x, " "))
    if(length(x) > 2) {
      x <- paste(x[2:3], collapse = "")
    } else {
      x <- x[2]
    }
    if (grepl("AM", x)) {
      temp <- unlist(strsplit(x, ""))
      temp <- paste(temp[1:2], collapse = "")
      if (temp != 12) {
         x <- gsub("AM", "", x)
         x <- paste(x, ":", "00", sep = "")
      } else {
        x <- gsub("AM", "", x)
        split_minhr <- unlist(strsplit(x, ":"))
        hours <- as.numeric(split_minhr[1])
        mins <- split_minhr[2]
        mil_hr <- hours - 12
        x <- paste(mil_hr, ":", mins, ":", "00", sep = "")
      }
    }
    if (grepl("PM", x)) {
      x <- gsub("PM", "", x)
      temp <- unlist(strsplit(x, ""))
      temp <- paste(temp[1:2], collapse = "")
      if (temp != "12") {
        split_minhr <- unlist(strsplit(x, ":"))
        hours <- as.numeric(split_minhr[1])
        mins <- split_minhr[2]
        mil_hr <- hours + 12
        x <- paste(mil_hr, ":", mins, ":", "00", sep = "")
      } else {
        x <- paste(x, ":", "00", sep = "")
      }
    }
  }
  if (grepl("/", x)) {
    x <- gsub("/", "", x)
    x <- unlist(strsplit(x, " "))
    x <- x[2]
  }
  x
}

```

## Use get_miltime function

```{r}
start_mil_time <- character(76)
end_mil_time <- character(76)
for (i in seq_len(length(sleep$Start.Time))) {
  start_mil_time[i] <- get_miltime(sleep$Start.Time[i])
}
for (i in seq_len(length(sleep$End.Time))) {
  end_mil_time[i] <- get_miltime(sleep$End.Time[i])
}
for(i in 60:76) {
  if (!is.na(start_mil_time[i])) {
    start_mil_time[i] <- paste(start_mil_time[i], ":", "00", sep = "")
  }
}
for(i in 60:76) {
  if (!is.na(end_mil_time[i])) {
    end_mil_time[i] <- paste(end_mil_time[i], ":", "00", sep = "")
  }
}

start_asTime <- as.numeric(chron(times = start_mil_time))
end_asTime <- as.numeric(chron(times = end_mil_time))
```

## Clean the times and dates and make final data frame

```{r, warning = FALSE}
sleep_final <- sleep[, c(1, 21, 15, 27, 29, 30)]
sleep_final$start_mil_time <- start_mil_time
sleep_final$end_mil_time <- end_mil_time
sleep_final$overall_score <- as.numeric(sleep_final$overall_score)
sleep_final$Number.of.Awakenings <- as.numeric(sleep_final$Number.of.Awakenings)
sleep_final$restlessness <- as.numeric(sleep_final$restlessness)
sleep_final$start_asTime <- start_asTime
sleep_final$end_asTime <- end_asTime

six_oclock <- as.numeric(chron(times = "18:00:00"))
sleep_final$start_asTime <- ifelse(start_asTime > .88, start_asTime - six_oclock,
                                   start_asTime + .25)
# Times greater than 12 AM: add a day (1) then subtract hours of 6 PM (18:00 equals .75 of
# a day)... i.e. add .25 of a day, or 6 hours, to start_asTime[i] to get time past 6 PM

sleep_final$end_asTime <- end_asTime + .25

# Convert end_asTime to hours past 6 PM for uniformity.

prob_rem <- with(sleep, as.numeric(Minutes.REM.Sleep)/as.numeric(Minutes.Asleep))
# Above produces warning that NAs are introduced by coercion, because there
# are NAs in these data.
sleep_final$prob_rem <- prob_rem
```

# Exploring the data

# Look for transformations needed

## Y variable

```{r}
sleep_final_final <- sleep_final[, c(1, 2, 3, 4, 9, 10, 11)]
sleep_lm <- lm(overall_score ~ . - Date, data = sleep_final_final)
require(car)
library(alr3)
summary(powerTransform(sleep_lm))
invResPlot(sleep_lm)
hist(sleep_final$overall_score, col = "dodgerblue2")
```

## X variables

```{r}
summary(powerTransform(as.matrix(sleep_final_final[, c(3:7)]) ~ 1))
```

## Do transforms

```{r}
sleep_final_final$overall_score <- sleep_final_final$overall_score ^ 3
```

# Fit the new model, output the summary

```{r}
sleep_lm_2 <- lm(overall_score ~ . - Date, data = sleep_final_final)
summary(sleep_lm_2)
```

# Check the diagnostics for validity/model weaknesses

```{r, warning = FALSE}
par(mfrow = c(2, 2))
plot(sleep_lm_2, col = "goldenrod", pch = 18, which = c(1:3))
n <- dim(sleep_final_final)[1]
plot(sleep_lm_2, col = "goldenrod", pch = 18, which = 5)
abline(v = 2 * (5 + 1) / n, lty = 2, col = "forestgreen", lwd = 2)
abline(h = c(2, -2), lty = 2, col = "dodgerblue", lwd = 2)
rect(xleft = 2 * (5 + 1) / n, xright = .3, ybottom = c(2, -3.5), ytop = c(3.5, -2),
     border = NULL, col = "gray60", density = 15)
```

### Bad Leverage Points and Influential Points

```{r}
plot(sleep_lm_2, which = 4, col = "goldenrod", lwd = 2)
hist(cooks.distance(sleep_lm_2), col = "dodgerblue3",
     main = "Distribution of Cook's Distances:\nWhich points, if removed, would change the regression model most?",
     xlab = "Cook's Distance",
     cex.main = .89)
```

